---
title: "Reproducible Research Course Project 1"
output:
  word_document: default
  html_document:
    Data_print: paged
    keep_md: yes
---
Data

The data for this assignment was downloaded from the course web site:

Dataset: Activity monitoring data [52K]
The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as NA)

date: The date on which the measurement was taken in YYYY-MM-DD format

interval: Identifier for the 5-minute interval in which measurement was taken

The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

###Loading and preprocessing the data
Download, unzip and load data into data frame data.

```{r, echo = TRUE}
knitr::opts_chunk$set(
  fig.path = "images/"
)
cls = c("integer", "character", "integer")
Data <- read.csv("activity.csv", head=TRUE, colClasses=cls, na.strings="NA")
head(Data)
```
###What is mean total number of steps taken per day?

```{r}
Data$date <- as.Date(Data$date)
Data_ign <- subset(Data, !is.na(Data$steps))
dailysum <- tapply(Data_ign$steps, Data_ign$date, sum, na.rm=TRUE, simplify=T)
dailysum <- dailysum[!is.na(dailysum)]

hist(x=dailysum,
     col="green",
     breaks=20,
     xlab="Daily total steps",
     ylab="Frequency",
     main="The distribution of daily total (missing data ignored)")
```

Mean and median of total number of steps taken per day

```{r, echo = TRUE}
mean(dailysum)
```

```{r, echo = TRUE}
median(dailysum)
```
###What is the average daily activity pattern?
```{r, echo = TRUE}
knitr::opts_chunk$set(
  fig.path = "images/"
)
int_avg <- tapply(Data_ign$steps, Data_ign$interval, mean, na.rm=TRUE, simplify=T)
Data_ia <- data.frame(interval=as.integer(names(int_avg)), avg=int_avg)

with(Data_ia,
     plot(interval,
          avg,
          type="l",
          xlab="5-min intervals",
          ylab="average steps in the interval across all days"))
```

```{r, echo = TRUE}
max_steps <- max(Data_ia$avg)
Data_ia[Data_ia$avg == max_steps, ]
```
###Imputing missing values
```{r, echo = TRUE}
sum(is.na(Data$steps))
```

```{r, echo = TRUE}
Data_impute <- Data
ndx <- is.na(Data_impute$steps)
int_avg <- tapply(Data_ign$steps, Data_ign$interval, mean, na.rm=TRUE, simplify=T)
Data_impute$steps[ndx] <- int_avg[as.character(Data_impute$interval[ndx])]
```

```{r, echo = TRUE}
knitr::opts_chunk$set(
  fig.path = "images/"
)
new_dailysum <- tapply(Data_impute$steps, Data_impute$date, sum, na.rm=TRUE, simplify=T)

hist(x=new_dailysum,
     col="blue",
     breaks=20,
     xlab="daily steps",
     ylab="frequency",
     main="The distribution of daily total (with missing data imputed)")
```

```{r, echo = TRUE}
mean(new_dailysum)
```

```{r,echo = TRUE}
median(new_dailysum)
```
###Are there differences in activity patterns between weekdays and weekends?
```{r,echo = TRUE}
is_weekday <- function(d) {
  wd <- weekdays(d)
  ifelse (wd == "Saturday" | wd == "Sunday", "weekend", "weekday")
}

wx <- sapply(Data_impute$date, is_weekday)
Data_impute$wk <- as.factor(wx)
head(Data_impute)
```

```{r,echo = TRUE}
knitr::opts_chunk$set(
  fig.path = "images/"
)
wk_Data <- aggregate(steps ~ wk+interval, data=Data_impute, FUN=mean)

library(lattice)
xyplot(steps ~ interval | factor(wk),
       layout = c(1, 2),
       xlab="Interval",
       ylab="Number of steps",
       type="l",
       lty=1,
       data=wk_Data)
```


